An Extensible Approach to Verification of Embedded Network Systems*

### 2014-04-25 tech-project/implementation.tex
working on memory allocation

\subsection{Memory Allocation}

@@

The \texttt{salloc()} is a first fit memory allocator. Rather than allocating on the heap, \texttt{salloc()} uses the stack. At compilation, the allocator provisions a fixed size of memory. It requires a small section of meta-data called a block which contains information of about the size of memory allocated, neighboring blocks, as well if the block is free or not. A sample representation of the stack shown in figure \ref{fig:stack}, is a typical example of allocated memory on the stack.

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/stack.pdf}
\caption{Representation of memory stack}
\end{figure}
\label{fig:stack}

@@

A memory allocator must provide a mechanism to release, or free memory in order to indicate that it is not being used and can be reallocated for something else. The \texttt{sfree()} operation is an analogous implementation to the standard C \texttt{free()} function for the stack.

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/sfree.pdf}
\caption{sfree to free memory}
\end{figure}
\label{fig:free}

\subsubsection{Optimizations}

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/fragmentation.pdf}
\caption{fragmentation}
\end{figure}
\label{fig:fragmentation}

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/split.pdf}
\caption{block splitting}
\end{figure}
\label{fig:split}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%
%           Memory Allocation Section              %
%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Memory Allocation}\label{sec:mem}

@@

The function \texttt{salloc()}, uses a first fit approach for allocation on the stack, rather than conventional, heap-based allocators. This approach requires that a fixed size of memory is chosen at compilation. In addition, \texttt{salloc()}, requires a section of meta-data, which is denoted as a \texttt{block}, for each record in the memory pool. A \texttt{block} holds referential information of neighboring blocks, the size of the record the block maintains, and if the block is free or is in use. Figure~\ref{fig:stack} is an example of the stack after several calls to \texttt{salloc()}.

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/stack.pdf}
\caption{Representation of memory stack}
\label{fig:stack}
\end{figure}

@@

A memory allocator must provide a mechanism to release, or free memory in order to indicate that it is not being used and can be reallocated. The \texttt{sfree()} function provides deallocation for memory allocated with \texttt{salloc()}. Figure~\ref{fig:free} shows an example set of memory before and after calling \texttt{sfree()}.

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/sfree.pdf}
\caption{sfree to free memory}
\label{fig:free}
\end{figure}

\subsubsection{Optimizing Memory Usage}

A common problem that can occur in memory allocation is fragmentation. During execution, a program can allocate and deallocate an arbitrary number of times. Figure~\ref{fig:fragmentation} shows problems that result from this. This problem is magnified on embedded systems due to their limited memory capacities. There are simple optimizations that can reduce fragmentation.

Block splitting, as shown in Figure~\ref{fig:split}, occurs during allocation. It allows blocks of greater size to be partitioned into the size requested by the allocator.

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/fragmentation.pdf}
\caption{fragmentation}
\label{fig:fragmentation}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[scale=.55]{figs/split.pdf}
\caption{block splitting}
\label{fig:split}
\end{figure}
